## Install & Load Packages

```{r}
# Load or activate them
library(terra)
library(sf)
library(geodata)
library(rnaturalearth)
```

## Get and Plot Vector and Raster Data

### Get country borders (Germany)

```{r}
ger <- ne_countries(country = "Kyrgyzstan", scale = "medium", returnclass = "sf")
plot(ger) # Plot German boundaries
```

### Get global temperature minimum data (WorldClim)

```{r}
clim <- geodata::worldclim_global(
  var = "tmin",
  res = 10,
  download = TRUE,
  path = "."
)
plot(clim) # Plot data
```

### Check if projections match, and reproject if needed

```{r}
ger.r <- st_transform(ger, st_crs(clim))
```

## Clip and Mask Data to Germany

### Crop precipitation (or temperature) data to extent of Germany

```{r}
clim.ger.crop <- terra::crop(clim, ger.r) # -> '::' because function exists in various packages
plot(clim.ger.crop) # Plot cropped result
```

### Mask data to exact shape of Germany

```{r}
clim.ger.mask <- terra::mask(clim.ger.crop, ger.r)
col_ramp <- colorRampPalette(c("#10abdaff", "#ff136aff"))(100)
plot(clim.ger.mask, col = col_ramp) # Plot masked result
```

# Extract Statistics for Country

## Extract the mean value (e.g., average temperature) for Germany

## Other statistics (e.g., "min", "max", "median") are also possible

```{r}
climGer.vect <- terra::extract(clim.ger.mask, ger, fun = mean)
```

## View extracted data

```{r}
climGer.vect # Plot the monthly mean values (columns 2–13) 
plot(unlist(climGer.vect[, 2:13]))
```
